---
title: "WUP_04_Alt"
author: "Cleo Ware"
date: "fall 2023"
output:
  html_document:
    df_print: paged
---

```{r include=FALSE,echo=FALSE}
require(knitr)
require(tidyverse)
require(tigerstats)
MYdata<-read.csv(file = "GSScsv.csv")

```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

The data is in the file GSScsv.csv and came from the 2018 General Social Survey.  The question to answer is to resolve this hypothesis.  
$H_0$ There is no difference between the average score on the WORDSUM test and the RACE of the person taking the test   
$H_a$ There is a significant difference in average score on the WORDSUM test between at least two different RACES

# Methods

<!--Mention that you intend to use chatGPT to the greatest extent possible. Give credit to chatGPT when you use it to create code in your report. Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

```{r eval=FALSE}
MYdata <- MYdata %>%
mutate(across(where(is.character), as.factor))
```

We have changed all the character variables into factor variables and we see that our problem is a type that needs a numerical variable called WORDSUM and a factor variable called RACE. We will be using group helper files to help us complete the analysis of numerical and factual data. 

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
basicCN <- ggplot(MYdata,aes(x=RACE,y=WORDSUM))

basicCN + geom_boxplot()
```

By looking at the above graph, there is no obvious difference between the different RACE scores.The WHITE RACE box is bigger, however we have no proof as to why.


### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
favstats(WORDSUM~RACE,data=MYdata)
```

By looking at the Means you can see that the BLACK and the WHITE RACES are close to each other. However, the OTHER RACE is different.

Due to it being smaller then BLACK and WHITE RACES it does show a difference.


## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
model1 <- aov(WORDSUM~RACE,data=MYdata)
summary.aov(model1)
```


```{r}
#we specify the margins starting at the bottom and proceeding clockwise - thus 8.7 is the left side margin to hold the names of the factors
par(mar = c(3, 8.7, 2, 2))
plot(TukeyHSD(model1),las=2)
```

Per the final plot,we can see there is no difference at zero between the OTHER-BLACK race and the WHITE-OTHER race.

The plot does show a difference between the White race and the OTHER race

White and other are the only races that appear to be different per the chart.

However, there was no difference between White and Black due to zero being present

# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

First I used my helper file Numerical~Cat.rmd to help me build a box plot.